---
title: "Quarter Review"
subtitle: "Week 8"
date: 2025-05-22
format: 
  html:
    embed-resources: true
toc: true
---


# Before we start

-   Questions regarding Final Paper or the class?

-   Next week no section, but office hours instead (Scott Hall 110)

# Agenda

-   Reviewing what we have learned

-   Model Interpretation and Diagnostics

-   Quarto formatting

# Quarter Review

[<i>{{< fa arrow-up-right-from-square >}}</i> View slides in new window](https://artur-baranov.github.io/nu-ps312-ds/slides/ps312-s_08.html){.btn .btn-primary role="button" target="_blank"}

<iframe height="500" width="100%" src="slides/ps312-s_08.html"></iframe>



```{r}
library(readxl)
cpds = read_excel("data/cpds.xlsx")
```

```{r}
#| warning: false
#| message: false
library(GGally)

ggpairs(cpds, columns = c("prefisc_gini", "eu", "openc", "poco"))
```

```{r}
cpds$eu = as.factor(cpds$eu)
cpds$poco = as.factor(cpds$poco)
```


```{r}
model_mlr = lm(prefisc_gini ~ as.factor(eu) + openc * as.factor(poco), cpds)
summary(model_mlr)
```

```{r}
library(ggeffects)

ggpredict(model_mlr, terms = c("openc", "poco")) |>
  plot()
```

```{r}
library(ggfortify)
autoplot(model_mlr, colour = "as.factor(poco)")
```


# What's next?

## Different Model Specifications

What if your outcome is not continuous?

| Outcome type | Model       | Functions                        |
|--------------|-------------|----------------------------------|
| Count data   | Poisson     | `glm()` with poisson link        |
| Binary       | Binomial    | `glm()` with binomial link       |
| Categorical  | Multinomial | `multinom()` from `nnet` library |

## Quarto Formatting



# Libraries and Functions covered

| Library      | Functions                          | Description                                          |
|--------------|------------------------------------|------------------------------------------------------|
| tidyverse    | `filter()`, `mutate()`, `ggplot()` | data wrangling and visualization                     |
| modelsummary | `modelsummary()`                   | present good looking tables                          |
| ggeffects    | `ggpredict()`                      | calculate and visualize marginal effects             |
| GGally       | `ggpairs()`, `ggcoef()`            | extension to ggplot                                  |
| ggfortify    | `autoplot()`                       | extension to ggplot for diagnostics                  |
| lmtest       | `bptest()`                         | statistical tests for diagnostics                    |
| car          | `vif()`                            | additional statistical tests for diagnostics         |

# Datasets covered

| Dataset                                           | Description                                | Link                                                     |
|---------------------------------------------------|--------------------------------------------|----------------------------------------------------------|
| **V-Dem**                                         | Measures democracy worldwide               | [V-Dem](https://www.v-dem.net/)                          |
| **World Happiness Report**                        | Annual happiness report                    | [World Happiness Report](https://worldhappiness.report/) |
| **Who Governs**                                   | Dataset on political elites                | [Who Governs](https://doi.org/10.1017/S0003055420000490) |
| **SIPRI**                                         | Data on military operations                | [SIPRI](https://www.sipri.org/databases)                 |
| **Comparative Political Data Set**                | A dataset covering political institutions  | [CPDS](https://www.cpds-data.org/)                       |
| 



# Check List

- [ ] I know that every lab has check list below! And I will use it to navigate what we have learned

- [ ] R doesn't scare me anymore

- [ ] I have developed the intuition behind application of quantitative methods
