---
title: "Regression Diagnostics"
subtitle: "Week 6"
date: 2025-05-08
format: 
  html:
    embed-resources: true
toc: true
---

# Before we start

-   Questions regarding Lab Assignment or the class?

<br><center>
<a href="https://artur-baranov.github.io/nu-ps312-ds/scripts/ps312-d_06.qmd" class="btn btn-primary" role="button" download="ps312-d_06.qmd" style="width:200px" target="_blank">Download script</a>
</center>
<br>
<center>
<a href="https://artur-baranov.github.io/nu-ps312-ds/data/appelloyle_jprdata.dta" class="btn btn-primary" role="button" download="appelloyle_jprdata.dta" style="width:200px" target="_blank">Download data</a>
</center><br>

# Agenda

-   Using dummy variables in the OLS

-   Introducing fixed effects and interactions

-   Discussing polynomials

## Set Up

Today we are working with WhoGov dataset. As usual, I recommend taking a look at their [codebook](https://politicscentre.nuffield.ox.ac.uk/media/4117/whogov_codebook.pdf).

```{r}
#| message: false

library(tidyverse)

whogov = read.csv("data/WhoGov.csv")
```

First of all, these are the following variables we are going to work with today:

-   `country_name` is a country name

-   `n_individuals` number of unique persons in the cabinet

-   `leaderexperience_continuous` the number of years the person has been leader of the country in total.

-   `leader_party` party of the leader

Start with exploring the distribution of number of unique persons in the cabinet (`n_individuals`)

```{r}
#| message: false

ggplot(whogov) +
  geom_histogram(aes(x = n_individuals)) 
```

Let’s examine whether a country's leader being independent from a political party is associated with having more or fewer members in their cabinet. First, let's create a dummy variable indicating if a leader is independent or non-independent. You can use 1 or 0 instead, but to make it more readable here we stick to more transparent labels.

```{r}
whogov = whogov %>%
  mutate(indep = ifelse(leader_party == "independent", "Independent", "Non-independent"))
```

Now, build a simple model and explore the effect. On average, being a non-independent (i.e. partisan) leader is associated with having 2.28 more members in their cabinet compared to independent leaders.

```{r}
lm(n_individuals ~ indep, whogov) %>%
  summary()
```

What if we want to know the effect relative to Non-independent leader? Let's `relevel()` the variable!

```{r}
#| eval: false

whogov$indep = relevel(whogov$indep, ref = "Non-independent")
```

Oops! This is why classes of data are important. Fix it!

```{r}
whogov$indep = as.factor(whogov$indep)
```

Now we can relevel the variable

```{r}
whogov$indep = relevel(whogov$indep, ref = "Non-independent")
```

Compare the models. Does the result sound reasonable? Pretty much. This is simply an inverse. But things get way more interesting if a categorical variable has more than 2 levels. You will see this later on. 

```{r}
lm(n_individuals ~ indep, whogov) %>%
  summary()
```




```{r}
mod = lm(log(n_individuals) ~ log(leaderexperience_continuous) * indep + country_name + as.factor(year), whogov)

library(lmtest)
plot(mod)
bptest(mod)

summary(mod)

```




